Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-27 WSSSPE1_report/WSSSPE1_summary_report.tex
reading and some small changes

\begin{abstract}
This technical report discusses the First Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE1).  The content
is based on the workshop itself, as documented in a set of
collaborative notes taken during the workshop, including discussion
about the keynote presentations, three themed panels that focus on
papers accepted by the workshop, other discussion, cross-cutting
issues, use cases, and conclusions.  It joins a nascent literature
seeking to understand what drives software work in science and how the
reward systems of science thereby shape the type of software work
undertaken, including the extent to which developers are motivated to
build software for the long-term, for the use of others, and whether
to work collaboratively or separately. Unique perspectives were
captured about issues such as documentation of APIs, software
deployment, and development practices; licenses on software; career
paths for scientific software writers; identifiers for software
through a mechanism such as a Digital Object Identifier or
publications of ``software papers''; drawing on evidence of software
contribution, and impact, recorded in online systems such as source
code repositories like GitHub.  For truly sustainable software there
should be no endpoint, as the software products continue to be used
and useful beyond the single institution, grant, and developer or
development team.

@@

Because progress in scientific research is dependent on the quality and
accessibility of software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments who are an integral and embedded part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
journal publication. An estimate of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences) and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 SIAM Annual Meeting on ``Reliable
Computational Science.''

@@

The workshop itself consisted of two keynote presentations and the
three panels/discussion sessions. The panels were organized based on a
classification of the workshop submissions into three categories,
following the themes of the call for papers as modified by the areas
of the submissions. Three to four representatives from each submission
category were appointed as panelists, and assigned to read a subset of
the paper in that category and then discuss them in the panel.

>>>>

\begin{abstract}
This technical report discusses the First Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE1).  The content
is based on the workshop itself, as documented in a set of
collaborative notes taken during the workshop, including discussion
about the keynote presentations, three themed panels that focus on
papers accepted by the workshop, other discussion, cross-cutting
issues, use cases, and conclusions.  It joins a nascent literature
seeking to understand what drives software work in science and how the
reward systems of science thereby shape the type of software work
undertaken, including the extent to which developers are motivated to
build software for the long-term, for the use of others, and whether
to work collaboratively or separately. Unique perspectives were
captured about issues such as documentation of APIs, software
deployment, and development practices; licenses on software; career
paths for scientific software developers; identifiers for software
through a mechanism such as a Digital Object Identifier or
publications of ``software papers''; drawing on evidence of software
contribution, and impact, recorded in online systems such as source
code repositories like GitHub. For truly sustainable software there
should be no endpoint, as the software products continue to be used
and useful beyond the single institution, grant, and developer or
development team.

@@

Because progress in scientific research is dependent on the quality and
accessibility of software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments who are an integral and embedded part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
journal publication \note{Ketan: we say report in abstract}. An estimate of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences) and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 SIAM Annual Meeting on ``Reliable
Computational Science.''

@@

The workshop itself consisted of two keynote presentations and the
three panels/discussion sessions. The panels were organized based on a
classification of the workshop submissions into three categories,
following the themes of the call for papers as modified by the areas
of the submissions. Three to four representatives from each submission
category were appointed as panelists, and assigned to read a subset of
the papers in that category and then discuss them in the panel.
